---
title: "Microbiome PD analysis"
author: "xxn"
date: "`r paste0('Initiated on 2019 August 22, compiled on ', format(Sys.time(), '%Y %b %d'))`"
output:
  html_document:
    chunk_output_type: console
    code_folding: hide
    editor_options: null
    fig_height: 8
    fig_width: 8
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
    indent: True
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F)
```

#After preprocessing of Fastq seq


```{r message=FALSE,echo=F,warning=F}
library(tidyverse)
library(phyloseq)
library(dada2)
library(DECIPHER)
library(phangorn)
library(gridExtra)
library(plotly)
library(ggplot2)
library(readxl)
library(kableExtra)
library(knitr)
library(janitor)
library(kableExtra)
library(knitr)
```

```{r}
load(file = "otu_0.RData")
set.seed(100)
dir_path <- "./FASTQ_Generation_2019-08-25_12_19_07Z-1341381/"
data_path <- list.files(dir_path)
file_path <- c()
fnFs <- c()
fnRs <- c()
sampleNames <- c()
for (i in 1:(length(data_path) - 1)){
  file_path[i] = paste0(dir_path,data_path[i],collapse = "/")
  fnFs0 <- sort(list.files(file_path[i], pattern="_R1_001.fastq"))
  fnRs0 <- sort(list.files(file_path[i], pattern="_R2_001.fastq")) 
  fnFs[i] <- file.path(file_path[i], fnFs0)
  fnRs[i] <- file.path(file_path[i], fnRs0)
  sampleNames[i] <- strsplit(fnFs0, "_")[[1]][1]
}

```

```{r}
seqtabNoC = res$OTU
taxTab = res$taxTab
fitGTR = res$fitGTR

```


```{r}
sample_sp = strsplit(sampleNames,"-t-")
sample_info = data.frame(id = 0,t = 0)
for (i in 1:length(sample_sp)){
  if(length(sample_sp[[i]]) == 2){
    id = sample_sp[[i]][1]
    id = sub(pattern = "-",replacement = "/",x = id)
    sample_time = sample_sp[[i]][2]
    if (sample_time == "minus2"){
      sample_time = -2
    }else if(is.na(as.numeric(sample_time))){
      sample_time = 0
    }else{
      sample_time = as.numeric(sample_time)
    }
    
  }else{
    id = sample_sp[[i]][1]
    sample_time = NA
  }
  sample_info = rbind(sample_info,c(id,sample_time))
}
sample_info = sample_info[-1,]
rownames(sample_info) = sampleNames
rownames(seqtabNoC) = sampleNames
```

```{r}
Sample.info = readxl::read_xlsx("../Project Database t=0 - de-identified.xlsx",sheet = 1)

idx = match(sample_info$id,Sample.info$`Neurogenetics ID`)

sample_info = cbind(sample_info,Sample.info[idx,c("Sex ID","Age","PD Status ID","Cohort Identifier ID","PD Age Dx")])
colnames(sample_info) = c("id","t","Sex","Age","PD","Cohort","PD_Age")

```

```{r}
ps <- phyloseq(otu_table(seqtabNoC, taxa_are_rows=FALSE), 
               sample_data(sample_info), 
               tax_table(taxTab),phy_tree(fitGTR$tree))
dim(seqtabNoC)
```

# Filter
```{r}
control = c("18-124-t-0","18-106-t-0","18-121-t-0","19-201-t-0","18-204-t-0","18-231-t-0","19-013-t-0","18-129-t-0","18-263-t-0","19-009-t-0")
ps <- subset_samples(ps, (!is.na(Age))&(!is.na(t))&(!is.na(PD))&(Cohort != "Longitudinal")&(t != 6)|(sample_names(ps) %in% control))
ps@sam_data$PD[is.na(ps@sam_data$PD)] = "no"
ps@sam_data$Cohort[is.na(ps@sam_data$Cohort)] = "Longitudinal"

```

In document, we got 88 samples, with 78 Parkinson’s disease patients and 10 controls. In this processing, we get total `r length(sample_names(ps))` samples, with 4 replicates, say, 18-033-t-0-rpt2, 18-033-t-0-rpt3, 18-069-t-0-rpt2, 18-069-t-0-rpt3.


```{r}
ps <- subset_taxa(ps, !is.na(Phylum) & !Phylum %in% c("", "uncharacterized"))

prevdf = apply(X = otu_table(ps),
               MARGIN = ifelse(taxa_are_rows(ps), yes = 1, no = 2),
               FUN = function(x){sum(x > 0)})

prevdf = data.frame(Prevalence = prevdf,
                    TotalAbundance = taxa_sums(ps),
                    tax_table(ps))
plyr::ddply(prevdf, "Phylum", function(df1){cbind(mean(df1$Prevalence),sum(df1$Prevalence))})
```

```{r}
filterPhyla = c("Elusimicrobia","Spirochaetes","Candidatus_Saccharibacteria","Cyanobacteria/Chloroplast")
# Filter entries with unidentified Phylum.
ps1 = subset_taxa(ps, !Phylum %in% filterPhyla)

prevdf1 = subset(prevdf, Phylum %in% get_taxa_unique(ps1, "Phylum"))
ggplot(prevdf1, aes(TotalAbundance, Prevalence / nsamples(ps),color=Phylum)) +
  # Include a guess for parameter
  geom_hline(yintercept = 0.05, alpha = 0.5, linetype = 2) +  geom_point(size = 2, alpha = 0.7) +
  scale_x_log10() +  xlab("Total Abundance") + ylab("Prevalence [Frac. Samples]") +
  facet_wrap(~Phylum) + theme(legend.position="none")
```

```{r}
prevalenceThreshold = floor(0.1 * nsamples(ps))
keepTaxa = rownames(prevdf1)[(prevdf1$Prevalence >= prevalenceThreshold)]
ps2 = prune_taxa(keepTaxa, ps)
dim(ps2@otu_table)
```

Filter out taxa with frequency less than 0.1 among all samples. 

# Unsupervised Result.

## PCoA of wUniFrac&BC among different PD state.
1. A quick PCoA using BC dissimilarity
$BC=\frac{\sum_{i=1}^{n}|X_{ij}-X_{jk}|}{\sum_{i=1}^n(X_{ij}+X_{jk})}$

```{r}
pslog <- transform_sample_counts(ps2, function(x) log(1 + x))
out.pcoa.log <- ordinate(pslog,  method = "MDS", distance = "bray")
evals <- out.pcoa.log$values[,1]
plot_ordination(pslog, out.pcoa.log, color = "t",
                  shape = "Cohort") +
  labs(col = "Time", shape = "Cohort")+
  coord_fixed(sqrt(evals[2] / evals[1])) + theme_classic()

```

```{r}

```

Transformation: $X'_{ij}=log(X_{ij}+1)$
Outlier: 18-203-t-2, most of its OTU is zero.

2. Principal coordinates analysis of weighted UniFrac distances using the relative abundance of OTUs among the patient groups. . Also Bray-Curtis dissimilarity analysis according to health status.

BC Dissimilarity.
```{r}
ps3 <- prune_samples(sample_names(ps2) != "18-203-t-2",ps2)
ps3ra <- transform_sample_counts(ps3, function(x){x / sum(x)})
#ps3ralog <- transform_sample_counts(ps3ra, function(x){log(x + eps)})

#ps3ralog <- microbiome::transform(ps3ra,"clr")
ps3ralog@sam_data$Cohort[ps3ralog@sam_data$Cohort == "Longitudinal"] = "Control"
out.pcoa3ra <- ordinate(ps3ralog,  method = "MDS", distance = "bray")
evals <- out.pcoa3ra$values[,1]
plot_ordination(ps3ralog, out.pcoa3ra, color = "t",
                  shape = "Cohort") +
  labs(col = "Time", shape = "Cohort")+
  coord_fixed(sqrt(evals[2] / evals[1])) + theme_classic()
```

wUniFrac Dissimilarity.  

$wUniFrac=\frac{\sum_{i=1}^nb_i|p_i^A-p_i^B|}{\sum_{i=1}^nb_i(p_i^A+p_i^B)}$

wUniFrac: Measure the dissimilarity from hierarchy clustering tree.  
$b_{i}$: length of branch i.
$p_i^A$: taxa proportions descending from the branch i for community A.  

```{r message=F,warning=F,echo=F}
out.pcoa3ra2 <- ordinate(ps3ra,  method = "MDS", distance = "wunifrac")
evals <- out.pcoa3ra2$values[,1]
plot_ordination(ps3ra, out.pcoa3ra2, color = "t",
                  shape = "Cohort") +
  labs(col = "Time", shape = "Cohort")+
  coord_fixed(sqrt(evals[2] / evals[1]))
```

phylogenetic tree is generated using "phangorn" package (Bayesian method), which incorporate background phylogenetic knowledge (Proir distribution) and data information(Sampling). (Take very long time....)


- After set high threshold (prevalence > 0.5, 67 variables left), still bad seperation in PCoA.

- Batch Effect(?) is not significant in this dataset.

```{r}
out.pcoa3ra$vectors[c("18-033-t-0-rpt2","18-033-t-0-rpt3","18-033-t-0"),1:2]
out.pcoa3ra$vectors[c("18-069-t-0-rpt2","18-069-t-0-rpt3","18-069-t-0"),1:2]
```


```{r}
ps3ra2 <- subset_samples(ps3ra, t %in% c(-2,0))

out.pcoa3ra3 <- ordinate(ps3ra2,  method = "MDS", distance = "bray")
evals <- out.pcoa3ra3$values[,1]
plot_ordination(ps3ra2, out.pcoa3ra3, color = "Cohort",
                  shape = "t") +
  labs(col = "Cohort", shape = "Time")+
  coord_fixed(sqrt(evals[2] / evals[1]))

```

Even if restrict t=0, it is bad seperation among three groups.

Consider the most extreme case: 
```{r}
ps3ra4 <- subset_samples(ps3ra, ((t == 0)&(PD == "no"))|(t == 4))

out.pcoa3ra4 <- ordinate(ps3ra4,  method = "MDS", distance = "bray")
evals <- out.pcoa3ra4$values[,1]
plot_ordination(ps3ra4, out.pcoa3ra4, color = "Cohort",
                  shape = "t") +
  labs(col = "Cohort", shape = "Time")+
  coord_fixed(sqrt(evals[2] / evals[1]))
```



## boxplot in different phylogenetic levels

```{r}
D = data.frame(rank = rank_names(ps3), number = 0)
for(i in 1:length(rank_names(ps3))){
  D[i,2] = length(table(ps3@tax_table[,i]))
}
D %>% kable %>% kable_styling()
```



boxplot in Phylum level.

```{r}
plot_abundance = function(physeq,title = "",
                          Facet = "Order", Color = "Phylum"){
  # Arbitrary subset, based on Phylum, for plotting
  p1f = subset_taxa(physeq, Phylum %in% c("Firmicutes"))
  mphyseq = psmelt(p1f)
  mphyseq <- subset(mphyseq, Abundance > 0)
  ggplot(data = mphyseq, mapping = aes_string(x = "sex",y = "Abundance",
                              color = Color, fill = Color)) +
    geom_violin(fill = NA) +
    geom_point(size = 1, alpha = 0.3,
               position = position_jitter(width = 0.3)) +
    facet_wrap(facets = Facet) + scale_y_log10()+
    theme(legend.position="none")
}
```

```{r}
eps <- 1e-5
p <- ncol(ps3@otu_table)
ps3ra3 <- transform_sample_counts(ps3ra2, function(x){(x + eps) / sum(x + p*eps)})
mphyseq1 <- psmelt(ps3ra2)
mphyseq2 <- psmelt(ps3ra3)
```

```{r}
ggplot(data = mphyseq2, mapping = aes_string(x = "Phylum",y = "Abundance",
                              color = "Cohort", fill = "Cohort")) +
    geom_boxplot() + scale_y_log10() + theme(axis.text.x = element_text(angle = 45)) + labs(title = "boxplot of Phylum with all abundance")

ggplot(data = mphyseq1, mapping = aes_string(x = "Phylum",y = "Abundance",
                              color = "Cohort", fill = "Cohort")) +
    geom_boxplot() + scale_y_log10() + theme(axis.text.x = element_text(angle = 45)) + labs(title = "boxplot of Phylum without 0 abundance")
```

A Ubiquituous Question: Is the zero the real one? 
- Technical zero
- Rounding zero
- Assignment zero(<50% assignment)


## density of alpha diversity
```{r}
library(vegan)
```

```{r}
S_shannon <- diversity(ps3@otu_table, index = "shannon", MARGIN = 1)
```

```{r}
S_simpson <- diversity(ps3@otu_table, index = "simpson", MARGIN = 1)
```

```{r}
D_alpha <- ps3@sam_data
D_alpha <- as.data.frame(as.matrix(D_alpha))
D_shannon <- D_alpha
D_shannon$shannon <- S_shannon
D_simpson <- D_alpha
D_simpson$simpson <- S_simpson

ggplot(data = D_shannon, mapping = aes_string(x = "t",y = "shannon",
                              color = "Cohort", fill = "Cohort")) +
    geom_boxplot() + theme(axis.text.x = element_text(angle = 45)) + labs(title = "boxplot of shannon alpha diversity") 


```
```{r}
ggplot(data = D_simpson, mapping = aes_string(x = "t",y = "simpson",
                              color = "Cohort", fill = "Cohort")) +
    geom_boxplot() + theme(axis.text.x = element_text(angle = 45)) + labs(title = "boxplot of simpson alpha diversity") 
```

```{r}
Illustration <- subset_samples(ps3ra,id == "18/205")
Illustration_melt <- psmelt(Illustration)
ps_phylum <- Illustration_melt %>% group_by(id,t,Phylum) %>% summarise(Abundance = sum(Abundance))
ggplot(ps_phylum,aes(x = t,y = Abundance,fill = Phylum)) + geom_bar(stat = "identity", width = 1) + scale_fill_viridis_d() + theme_classic()
```

